---
title: "BT4015 Project"
author: "Group 1"
date: "2022-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setting up

```{r pressure, echo=FALSE, warning=FALSE, error=FALSE}
library(GISTools)
library(tmap)
library(sp)
library(sf)
library(raster)
library(rgdal)
library(spatstat)
library(rgeos)
```

## Set seed

```{r}
set.seed(42)
```

## Loading data

### Polygons and lines

```{r, warning=FALSE, error=FALSE}
planning_area_sf <- st_read("data/town-councils/Planning_Area_Census2010.kml")
sg <- readOGR("./data/town-councils", "Region_Census2010")
# set_RGEOS_CheckValidity(2L)
# sg_combined <- gUnaryUnion(sg)
```

### Points

Next we read in point data. The first set of point data is the Park Connector (PCN) access points themselves. Next, we have data on various public amenities like supermarkets, hawker centres, pharmacies as well as transportation infrastructure like MRT exits.

```{r}
access_points_sf <- st_read("data/PCNs/PCN_Access_Points.kml")
supermarkets_sf <- st_read("data/amenities/supermarkets/supermarkets.kml")
hawkercentres_sf <- st_read("data/amenities/hawker-centres/hawkercentre.kml")
mrt_exits_sf <- st_read("data/MRTs/lta-mrt-station-exit-kml.kml")
pharmacy_sf <- st_read("data/amenities/pharmacies/retail-pharmacy-locations-kml.kml")
parks <- st_read("data/parks/parks-geojson.geojson")
busstops_sf <- st_read("data/bus-stops/BusStop.shp")
bike_racks <- st_read("data/lta-bicycle-rack/lta-bicycle-rack-geojson.geojson")


class(access_points_sf)
class(supermarkets_sf)
class(hawkercentres_sf)
class(mrt_exits_sf)
class(pharmacy_sf)
class(parks)
class(busstops_sf)
class(bike_racks)
```

### Raster

Next, we read in raster data, which consists of only population density data.

```{r}
GDALinfo("data/population-density/popmap15adj.tif")
```

```{r}
pop_density <- raster("data/population-density/popmap15adj.tif")
pop_density
```

# Exploratory Data Analysis

## Park Connectors

First, we plot the PCN access points. On first glance, it is apparent that these points cannot be randomly distributed, with noticeable clusters in the Choa Chu Kang/Bukit Batok area as well as the Woodlands/Sembawang area. The points also seem to occur as a "string" of points, which obviously stands to reason since these are access points to Park Connectors which are essentially lines.

```{r}
tmap_mode("view")
tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) + tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red")+ tm_layout(title="PCN Access Points in Singapore", title.size = 0.65, title.position = c(0.02, "bottom"))
```

We also want to get a buffer of 500m around the PCN access points

```{r}
access_points_500buf <- st_buffer(access_points_sf, dist = 500)
access_points_500buf <- st_union(access_points_500buf)
access_points_500buf_sp <- as_Spatial(access_points_500buf)
```

## Plot various amenities over the PCN access points
Now, we will display the geographical distribution of the various amenities (supermarkets, hawker centres, etc.) on top of the locations of the PCN access points.

### Supermarkets 
We plot the locations of supermarkets overlaid with the access points on a map of Singapore. We see that most access points are located very close to supermarkets.
#### Plot just the points
```{r}
tmap_mode("view")

tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) +
  tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red", legend.col.show=TRUE, shapes.labels='PCAPs') + tm_add_legend(type="symbol", labels=c("PCAPs"), col='red') +
  tm_shape(supermarkets_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "skyblue", legend.col.show=TRUE, shapes.labels='supermarkets') + tm_add_legend(type="symbol", labels=c("supermarkets"), col='skyblue') + 
  tm_shape(access_points_500buf) + tm_borders("brown") +
  tm_layout(title="PCAPs and supermarkets", title.size = 0.65, title.position = c(0.02, "bottom"), legend.show=TRUE, legend.outside=FALSE, legend.text.size=0.7) 
```

#### calculate % of PCAPs within 500m of a supermarket
```{r}
supermarkets_500buf <- st_buffer(supermarkets_sf, dist = 500)
supermarkets_500buf <- st_union(supermarkets_500buf)

supermarkets_500buf_sf <- st_sf(supermarkets_500buf)
supermarkets_500buf_sp <- as_Spatial(supermarkets_500buf_sf)
access_points_sp <- as_Spatial(access_points_sf)

# count number of PCAPs within 500m of a supermarket
PCAPs_near_supermarkets <- poly.counts(access_points_sp, supermarkets_500buf_sp)

print(PCAPs_near_supermarkets)
print(dim(access_points_sf)[1])
print(PCAPs_near_supermarkets / dim(access_points_sf)[1])
```

### Hawker centres
For hawker centres, the correlation is not as obvious.
```{r}
tmap_mode("plot")

tm_shape(sg)+ tm_borders() + tmap_options(check.and.fix = TRUE) + tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red") + tm_add_legend(type="symbol", labels=c("PCAPs"), col='red') +
  tm_shape(hawkercentres_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "green") + tm_add_legend(type="symbol", labels=c("hawker centres"), col='green') + 
  tm_shape(access_points_500buf) + tm_borders("brown") +
  tm_layout(title="PCAPs and Hawker Centres", title.size = 0.65, title.position = c(0.02, "bottom"))
```

#### calculate % of PCAPs within 500m of a hawker centre
```{r}
hawkercentres_500buf <- st_buffer(hawkercentres_sf, dist = 500)
hawkercentres_500buf <- st_union(hawkercentres_500buf)

hawkercentres_500buf_sf <- st_sf(hawkercentres_500buf)
hawkercentres_500buf_sp <- as_Spatial(hawkercentres_500buf_sf)
access_points_sp <- as_Spatial(access_points_sf)

# count number of PCAPs within 500m of a supermarket
PCAPs_near_hawkercentres <- poly.counts(access_points_sp, hawkercentres_500buf_sp)

print(PCAPs_near_hawkercentres)
print(dim(access_points_sf)[1])
print(PCAPs_near_hawkercentres / dim(access_points_sf)[1])
```

### MRT exits
For MRT exits, we notice that the Downtown Core area has a high concentration of exits. This could be because many of the stations in that area have many exits, sometimes as many as 7 or 8. This makes it seem more numerous. We can try getting a dataset of just MRT stations themselves to help us narrow down our investigation. Unlike the MRT exits, we see that there are very few PCN access points in the Downtown Core area. This could be because of the lower priority assigned to cycling paths over mass rapid transit in a space-scarce area like the Downtown Core.
```{r}
tmap_mode("plot")

tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) + 
  tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red") + tm_add_legend(type="symbol", labels=c("PCAPs"), col='red') +
  tm_shape(mrt_exits_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "navyblue") + tm_add_legend(type="symbol", labels=c("MRT exits"), col='navyblue') +
  tm_shape(access_points_500buf) + tm_borders("brown") +
  tm_layout(title="PCAPs and MRT exits", title.size = 0.65, title.position = c(0.02, "bottom"))
```



#### calculate % of PCAPs within 500m of an MRT exit
```{r}
mrt_exits_500buf <- st_buffer(mrt_exits_sf, dist = 500)
mrt_exits_500buf <- st_union(mrt_exits_500buf)

mrt_exits_500buf_sf <- st_sf(mrt_exits_500buf)
mrt_exits_500buf_sp <- as_Spatial(mrt_exits_500buf_sf)
access_points_sp <- as_Spatial(access_points_sf)

# count number of PCAPs within 500m of a supermarket
PCAPs_near_mrt_exits <- poly.counts(access_points_sp, mrt_exits_500buf_sp)

print(PCAPs_near_mrt_exits)
print(dim(access_points_sf)[1])
print(PCAPs_near_mrt_exits / dim(access_points_sf)[1])
```

### Bus stops
```{r}
tmap_mode("plot")

tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) + 
  tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red") + tm_add_legend(type="symbol", labels=c("PCAPs"), col='red') +
  tm_shape(busstops_sf) + tm_symbols(size = 0.1, alpha = 0.2, col = "blue") + tm_add_legend(type="symbol", labels=c("bus stops"), col='blue') +
  tm_shape(access_points_500buf) + tm_borders("brown") + 
  tm_layout(title="PCAPs and bus stops", title.size = 0.65, title.position = c(0.02, "bottom"))
```

```{r}
busstops_sp <- as_Spatial(busstops_sf)
busstops_sp <- spTransform(busstops_sp, crs(access_points_500buf_sp))
```

#### calculate % of PCAPs within 500m of a bus stop
```{r}
busstops_500buf <- st_buffer(busstops_sf, dist = 500)
busstops_500buf <- st_union(busstops_500buf)

busstops_500buf_sf <- st_sf(busstops_500buf)
busstops_500buf_sp <- as_Spatial(busstops_500buf_sf)
busstops_500buf_sp <- spTransform(busstops_500buf_sp, crs(access_points_sp))

# count number of PCAPs within 500m of a supermarket
PCAPs_near_busstops <- poly.counts(access_points_sp, busstops_500buf_sp)

print(PCAPs_near_busstops)
print(dim(access_points_sf)[1])
print(PCAPs_near_busstops / dim(access_points_sf)[1])
```

### Parks
```{r}
tmap_mode("plot")

tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) + 
  tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red") + tm_add_legend(type="symbol", labels=c("PCAPs"), col='red') +
  tm_shape(parks) + tm_symbols(size = 0.1, alpha = 0.5, col = "grey") + tm_add_legend(type="symbol", labels=c("parks"), col='grey') +
  tm_shape(access_points_500buf) + tm_borders("brown") + 
  tm_layout(title="PCAPs and parks", title.size = 0.65, title.position = c(0.02, "bottom"))
```

```{r}
park_sp <- as_Spatial(parks)
```

#### calculate % of PCAPs within 500m of a park
```{r}
parks_500buf <- st_buffer(parks, dist = 500)
parks_500buf <- st_union(parks_500buf)
parks_500buf_sp <- as_Spatial(parks_500buf)
PCAPs_near_parks <- poly.counts(access_points_sp, parks_500buf_sp)
print(PCAPs_near_parks / dim(access_points_sf)[1])
```

### Bicycle racks
```{r}
tmap_mode("plot")

tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) + 
  tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "red") + tm_add_legend(type="symbol", labels=c("PCAPs"), col='red') +
  tm_shape(bike_racks) + tm_symbols(size = 0.1, alpha = 0.5, col = "yellow") + tm_add_legend(type="symbol", labels=c("bicycle racks"), col='yellow') +
  tm_shape(access_points_500buf) + tm_borders("brown") + 
  tm_layout(title="PCAPs and bicycle racks", title.size = 0.65, title.position = c(0.02, "bottom"))
```

#### calculate % of PCAPs within 500m of a bicycle racks
```{r}
bikeracks_500buf <- st_buffer(bike_racks, dist = 500)
bikeracks_500buf <- st_union(bikeracks_500buf)
bikeracks_500buf <- as_Spatial(bikeracks_500buf)
access_points_sp <- as_Spatial(access_points_sf)

PCAPs_near_bikeracks <- poly.counts(access_points_sp, bikeracks_500buf)
print(PCAPs_near_bikeracks / dim(access_points_sf)[1])
```

### Population density
Interestingly, areas with the highest population density (Serangoon, Hougang and the areas surrounding Paya Lebar) seem to be relatively underserved by acces points as compared to Western regions like Bukit Batok which has a much lower population density.

Standing out as well is the central part of Singapore, which has virtually zero population density. This region encompasses the MacRitchie Reservoir area and naturally is not home to any human population density. It might be useful to exclude this region in later analysis to prevent confounding.
```{r}
tmap_mode("plot")

tm_shape(pop_density, bbox=st_bbox(sg)) + tm_raster(title="Population Density") +
  tm_shape(access_points_sf) + tm_symbols(size = 0.1, alpha = 0.5, col = "green") + tm_add_legend(type="symbol", labels=c("PCAPs"), col='green') +
  tm_shape(sg) + tm_borders() + 
  tm_layout(title="PCAPs and Population Density", title.size = 0.65, title.position = c(0.02, "bottom"), legend.position=c("right", "bottom"))
```

# Spatial Analysis
## Global Density
### Extracting planning areas to observe
```{r, echo=FALSE, }
training_sf <- planning_area_sf[planning_area_sf$Name %in% c("YISHUN","JURONG EAST", "JURONG WEST", "TAMPINES", "PASIR RIS", "BEDOK"),]
```

### Calculating global density
```{r}
sg_area <- 728.6 #km2
supermarkets_gd <- dim(supermarkets_sf)[1] / sg_area
hawkercentres_gd <- dim(hawkercentres_sf)[1] / sg_area
mrt_exits_gd <- dim(mrt_exits_sf)[1]/ sg_area
pharmacy_gd <- dim(pharmacy_sf)[1]/ sg_area
```

## Local Density Analysis
### Quadrat Density
get extent of Singapore
```{r}
sg <- readOGR("./data/town-councils", "Region_Census2010")
sg
sg_window <- as.owin(sg)
```

```{r}
# convert crs of access_points_sf to crs of sg
access_points_q <- st_transform(access_points_sf, crs(sg))
# convert sf to ppp
access_points_ppp <- as.ppp(st_coordinates(access_points_q), W=sg_window)
```

Try plotting quadrats of various sizes.
```{r}
Q6 <- quadratcount(access_points_ppp, nx=6, ny=6)
plot(Q6)
Q20 <- quadratcount(access_points_ppp, nx=20, ny=20)
plot(Q20)
```

From the quadrat density graph, we again see that most access points are concentrated in the North and West.
```{r}
# plot the density raster
plot(intensity(Q20, image=TRUE), main=NULL, las=1)
# plot access points
plot(access_points_ppp, pch=20, cex=0.6, col=rgb(0,0,0,.5), add=TRUE)
```

### KDE functions
```{r}
library(jsonlite)
library(splitstackshape)
library(sp)
library(leaflet)
library(data.table)
library(KernSmooth)
library(maptools)
library(rgdal)

kde_jsonlist_data<-function(list){
  dt<-data.table()
  for (str in list){
    data = fromJSON(str)
    latlng = data.frame(ll=data[["SrchResults"]][["LatLng"]])
    data_coords<-data.frame(cSplit(latlng,'ll',','))[-1,] 
    data_coords <- data_coords[, c(2, 1)]
    setDT(data_coords)
    dt <-rbindlist(list(dt,data_coords), use.names=TRUE)
  }
  kde_display(dt)
}

kde_sf_data<-function(sf){
  sf_coords<-st_coordinates(sf)
  kde_display(sf_coords)
}

##Renders heatmap based on calculated KDE
kde_display<-function(arr){
  kde <- bkde2D(arr,bandwidth=c(0.005, 0.003), gridsize = c(1000,1000))
  KernelDensityRaster <- raster(list(x=kde$x1 ,y=kde$x2 ,z = kde$fhat))
  KernelDensityRaster@data@values[which(KernelDensityRaster@data@values < 30)] <- NA
  palRaster <- colorNumeric("Spectral", domain = KernelDensityRaster@data@values, na.color = "transparent")
  
  leaflet() %>% addTiles() %>% 
    addRasterImage(KernelDensityRaster, 
                   colors = palRaster, 
                   opacity = .8) %>%
    addLegend(pal = palRaster, 
              values = KernelDensityRaster@data@values, 
              title = "Kernel Density")
}
```

#### Plotting KDEs for ESDA
```{r}
kde_sf_data(access_points_sf)
```

```{r}
kde_sf_data(supermarkets_sf)
```

```{r}
kde_sf_data(hawkercentres_sf)
```

```{r}
# amenities_list<-c("data\\amenities\\others\\exercisefacilities.json","data\\amenities\\others\\libraries.json","data\\amenities\\others\\preschools_location.json","data\\amenities\\others\\communityclubs.json")
amenities_list<-c("data/amenities/others/exercisefacilities.json","data/amenities/others/libraries.json","data/amenities/others/preschools_location.json","data/amenities/others/communityclubs.json")
kde_jsonlist_data(amenities_list)
```

#### Quadrat density - based on population density
```{r}
pop_density_q <- projectRaster(pop_density, crs = crs(sg))

brks <- c(0, 50, 100, 150, 200, 250, 300)
zcut <- cut(pop_density_q, breaks = brks)
pop_tess <- tess(image = zcut)

Qpop <- quadratcount(access_points_ppp, tess = pop_tess)
```

```{r}
plot(Qpop, main="", las=1)
```

```{r}
# compute density within each quadrat
Qpop_density <- intensity(Qpop)
Qpop_density
```

```{r}
cl <-  interp.colours(c("blue", "green", "yellow", "orange", "red", "darkred"), 6)
# plot the density raster
plot(intensity(Qpop, image=TRUE), main=NULL, las=1, col=cl)
# plot access points
plot(access_points_ppp, pch=20, cex=0.6, col=rgb(0,0,0,.5), add=TRUE)
```

## Density-based Analysis (Monte Carlo)
### get baseline ANN value
```{r}
ann_observed <- mean(nndist(access_points_ppp, k=1))
ann_observed
# observed ANN distance when k=2
ann_observed_k2 <- mean(nndist(access_points_ppp, k=2))
```
### Define function to make quadrat
```{r}
sf_to_quad_dens <- function(amenity_sf, sg_crs=crs(sg), window=sg_window) {
  # convert crs of access_points_sf to crs of sg
  amenity_q <- st_transform(amenity_sf, sg_crs)
  # convert sf to ppp
  amenity_ppp <- as.ppp(st_coordinates(amenity_q), W=window)
  
  amenity_Qpop <- quadratcount(amenity_ppp, nx=20, ny=20)
  plot(amenity_Qpop)
  amenity_density <- intensity(amenity_Qpop, image=TRUE)
  plot(amenity_density)
  print(class(amenity_density))
  return(amenity_density)
}
```

```{r}
sf_to_quad_dens(mrt_exits_sf)
```

### Define function to run Monte Carlo and print ANN distribution
```{r}
runMC <- function(nsims, npoints, f, ann_observed, window=sg_window) {
  ann <- vector(length = nsims) # empty object to be used to store simulated ANN values
  if (class(f)=="RasterLayer") {
    f_im <- as.im(f)
  } else {
    if (class(f) == "im") {
      f_im = f
    } else {
      f_im <- as.im(f, W=window)
    }
  }
  for (i in 1:nsims) {
    random_point <- rpoint(n=npoints, f=f_im)
    ann[i] <- mean(nndist(random_point, k=1))
  }
  
  # plot the last simulation
  Window(random_point) <- sg_window
  plot(f_im)
  plot(random_point, pch=16, main=NULL, cols=rgb(0,0,0,0.5), add=TRUE)
  
  # print statistics
  pval <- sum(ann_observed > ann)/nsims
  mean_ann <- mean(ann)
  sd_ann <- sd(ann)
  print(paste('p-value: ', pval))
  print(paste('mean ANN dist: ', mean_ann))
  print(paste('sd of ANN dist: ', sd_ann))
  
  # plot histogram
  hist(ann, main=NULL, las=1, breaks=40, col="bisque", xlim=range(ann_observed, ann))
  abline(v=ann_observed, col="blue")
}
```


### Population density
```{r}
pop_density_im <- as.im(pop_density_q)
runMC(1000, dim(access_points_sf)[1], pop_density_q, ann_observed)
```

### MRT exits
```{r}
# convert crs of access_points_sf to crs of sg
mrt_exits_q <- st_transform(mrt_exits_sf, crs(sg))
# convert sf to ppp
mrt_exits_ppp <- as.ppp(st_coordinates(mrt_exits_q), W=sg_window)

MRT_Qpop <- quadratcount(mrt_exits_ppp, nx=20, ny=20)
plot(MRT_Qpop)
```

```{r}
MRT_density <- intensity(MRT_Qpop, image=TRUE)
plot(MRT_density)
class(MRT_density)
```

perform Monte Carlo simulation using MRT density as probabilities
```{r}
runMC(1000, dim(access_points_sf)[1], MRT_density, ann_observed)
```

### Bus stops
```{r}
# convert crs of access_points_sf to crs of sg
busstops_q <- st_transform(busstops_sf, crs(sg))
# convert sf to ppp
busstops_ppp <- as.ppp(st_coordinates(busstops_q), W=sg_window)

busstop_Qpop <- quadratcount(busstops_ppp, nx=20, ny=20)
plot(busstop_Qpop)
```

```{r}
busstop_density <- intensity(busstop_Qpop, image=TRUE)
plot(busstop_density)
class(busstop_density)
```

```{r}
runMC(1000, dim(access_points_sf)[1], busstop_density, ann_observed)
```

## Distance-based Analysis (Monte Carlo)
### MRT Exits
Create a raster using the extent of Singapore
```{r}
# r <- raster(ncol=1000, nrow=1000, xmn= 2637.869, xmx=56396.44, ymn=15748.72, ymx=50256.33)
r <- raster(ncol=1000, nrow=1000, xmn= sg@bbox['x','min'], xmx=sg@bbox['x','max'], ymn=sg@bbox['y','min'], ymx=sg@bbox['y','max'])
```

Convert MRT exits with the same CRS as sg to an sp object
```{r}
mrt_exits_q_sp <- as(mrt_exits_q, "Spatial")
class(mrt_exits_q_sp)
```

Get x and y coordinates from MRT exits
```{r}
distances_mrt <- distanceFromPoints(object = r, xy=coordinates(mrt_exits_q_sp)[,1:2])
plot(distances_mrt)
```

Mask the distances_mrt raster with the shape of Singapore
```{r echo=FALSE, error=FALSE}
distances_mrt <- mask(distances_mrt, sg)
plot(distances_mrt)
plot(access_points_ppp, add = TRUE)
```

Convert distances_mrt from raster to im
```{r}
class(distances_mrt)
distances_mrt_im <- as.im(distances_mrt)
class(distances_mrt_im)
```

Run Monte Carlo simulation for 1000 epochs using distance raster to generate probabilities
```{r}
runMC(1000, dim(access_points_sf)[1], distances_mrt, ann_observed)
```


We notice that the points are generated mostly far away from the MRT exits (white = closest to MRT exits).
This calls for the need to invert the distance raster in some way before feeding it as an underlying function for the Monte Carlo to use to generate probabilities.

### Need to invert the distance raster
We need to figure out a function to invert the values in the raster. 
Also, we have to account for the fact that the f argument in the rpoint function only takes positive values to generate probabilities.
##### Try inverting using e\^(-x)
```{r}
runMC(1000, dim(access_points_sf)[1], exp(-distances_mrt_im), ann_observed)
```


The exponential function places too much weight on areas with very small distances.

##### Try inverting using reciprocal -\> 1/x
```{r}
runMC(1000, dim(access_points_sf)[1], 1/distances_mrt_im, ann_observed)
```

```{r}
n <- 1000 # number of simulations
ann_distances_mrt_recip <- vector(length = n) # empty object to be used to store simulated ANN values

for (i in 1:n) {
  random_point <- rpoint(n=access_points_ppp$n, f=1**(-distances_mrt_im))
  ann_distances_mrt_recip[i] <- mean(nndist(random_point, k=1))
}

Window(random_point) <- sg_window
plot(distances_bikeracks)
plot(random_point, pch=16, main=NULL, cols=rgb(0,0,0,0.5), add=TRUE)
```

##### try with flipping -\> y = -x + a where a = max(xi)
```{r}
inv_distances_mrt_im = -distances_mrt_im + max(distances_mrt_im)

runMC(1000, dim(access_points_sf)[1], inv_distances_mrt_im, ann_observed)
```



### Parks
```{r}
parks2 <- st_transform(parks, crs(sg))
```

```{r}
parks_sp <- as(parks2, "Spatial")
class(parks_sp)
```

```{r}
coordinates(parks_sp)[,1:2]
```

Get x and y coordinates from parks

```{r}
distances_parks <- distanceFromPoints(object = r, xy=coordinates(parks_sp)[,1:2])
plot(distances_parks)
```

Mask the distances_parks raster with the shape of Singapore

```{r}
distances_parks <- mask(distances_parks, sg)
plot(distances_parks)
plot(access_points_ppp, add = TRUE)
```

Convert distances_parks from raster to im

```{r}
class(distances_parks)
distances_parks_im <- as.im(distances_parks)
class(distances_parks)
```

Run Monte Carlo simulation for 1000 epochs
```{r}
inv_distances_parks_im = -distances_parks_im + max(distances_parks_im)
runMC(1000, dim(access_points_sf)[1], inv_distances_parks_im, ann_observed)
```


### Bus stops


```{r}
busstops_mc <- spTransform(busstops_sp, crs(sg))
```

```{r}
class(busstops_mc)
coordinates(busstops_mc)
```

Get x and y coordinates from parks

```{r}
distances_busstops <- distanceFromPoints(object = r, xy=coordinates(busstops_mc))
plot(distances_busstops)
```

Mask the distances_parks raster with the shape of Singapore

```{r}
distances_busstops <- mask(distances_busstops, sg)
plot(distances_parks)
plot(access_points_ppp, add = TRUE)
```

Convert distances_parks from raster to im

```{r}
class(distances_busstops)
distances_busstops_im <- as.im(distances_busstops)
class(distances_busstops_im)
```

Run monte carlo simulation for 1000 epochs
```{r}
inv_distances_busstops_im <- -distances_busstops_im + max(distances_busstops_im)
runMC(1000, dim(access_points_sf)[1], inv_distances_busstops_im, ann_observed)
```

### Bicycle racks

```{r}
bike_racks_sp <- as_Spatial(bike_racks)
bikeracks_mc <- spTransform(bike_racks_sp, crs(sg))
```

```{r}
class(bikeracks_mc)
coordinates(bikeracks_mc)[,1:2]
```

Get x and y coordinates from parks

```{r}
distances_bikeracks <- distanceFromPoints(object = r, xy=coordinates(bikeracks_mc)[,1:2])
plot(distances_bikeracks)
```

Mask the distances_bikeracks raster with the shape of Singapore

```{r}
distances_bikeracks <- mask(distances_bikeracks, sg)
plot(distances_bikeracks)
plot(access_points_ppp, add = TRUE)
```

Convert distances_parks from raster to im

```{r}
class(distances_bikeracks)
distances_bikeracks_im <- as.im(distances_bikeracks)
class(distances_bikeracks_im)
```

Run monte carlo simulation for 1000 epochs
```{r}
inv_distances_bikeracks_im <- -distances_bikeracks_im + max(distances_bikeracks_im)
runMC(1000, dim(access_points_sf)[1], inv_distances_bikeracks_im, ann_observed)
```


#Regional analysis
## Geylang
```{r}
sf_use_s2(FALSE)
regional_sf <- planning_area_sf[planning_area_sf$Name %in% c("GEYLANG"),]
ap_regional_sf<-st_filter(access_points_sf,regional_sf)
mrt_regional_sf<-st_filter(mrt_exits_sf,regional_sf)
bikeracks_regional_sf<-st_filter(bike_racks,regional_sf)
parks_regional_sf<-st_filter(parks,regional_sf)

regional_st<-st_transform(regional_sf,crs(regional_sf))
regional_sf <- st_transform(regional_sf, crs = 6345)
regional_sp <- as_Spatial(st_zm(regional_sf))
regional_window <- as.owin(regional_sp)

```

### Regional Access Points data
```{r}
ap_regional_q <- st_transform(ap_regional_sf, crs(regional_sp))
ap_regional_ppp <- as.ppp(st_coordinates(ap_regional_q), W=regional_window)

ap_regional_Q20 <- quadratcount(ap_regional_ppp, nx=20, ny=20)
ap_regional_distribution <- intensity(ap_regional_Q20, image=TRUE)

ann_observed <- mean(nndist(ap_regional_ppp, k=1))
ann_observed
```

### Regional Amenity Density-based Monte Carlo
```{r}
amenities <- list(mrt_regional_sf, bikeracks_regional_sf, parks_regional_sf)
for (i in 1:length(amenities)) {
  amenity_regional_sf <- amenities[[i]]
  amenity_regional_q <- st_transform(amenity_regional_sf, crs(regional_sp))
  amenity_regional_ppp <- as.ppp(st_coordinates(amenity_regional_q), W=regional_window)
  
  amenity_regional_Q20 <- quadratcount(amenity_regional_ppp, nx=20, ny=20)
  amenity_regional_distribution <- intensity(amenity_regional_Q20, image=TRUE)
  
  runMC(1000, ap_regional_ppp$n, amenity_regional_distribution, ann_observed)
}
```

### Regional Amenity Distance-based Monte Carlo
```{r}
r <- raster(ncol=1000, nrow=1000, xmn= regional_sp@bbox['x','min'], xmx=regional_sp@bbox['x','max'], ymn=regional_sp@bbox['y','min'], ymx=regional_sp@bbox['y','max'])

amenities <- list(mrt_regional_sf, parks_regional_sf, bikeracks_regional_sf)
for (i in 1:length(amenities)) {
  amenity_regional_sf <- amenities[[i]]
  amenity_regional_q = st_transform(amenity_regional_sf, crs(regional_sp))
  #Convert MRT exits with the same CRS as sg to an sp object
  amenity_regional_q_sp <- as(amenity_regional_q, "Spatial")
  
  #Get x and y coordinates from MRT exits
  amenity_regional_distances <- distanceFromPoints(object = r, xy=coordinates(amenity_regional_q_sp)[,1:2])
  plot(amenity_regional_distances)
  
  #Mask the distances_mrt raster with the shape of Singapore
  amenity_regional_distances <- mask(amenity_regional_distances, regional_sp)
  
  #Convert distances_mrt from raster to im
  amenity_regional_distances_distribution <- as.im(amenity_regional_distances)
  inv_amenity_regional_distances_distribution = -amenity_regional_distances_distribution + max(amenity_regional_distances_distribution)
  
  runMC(1000, ap_regional_ppp$n, inv_amenity_regional_distances_distribution, ann_observed)
}

```

## Jurong West
```{r}
sf_use_s2(FALSE)
regional_sf <- planning_area_sf[planning_area_sf$Name %in% c("JURONG WEST"),]
ap_regional_sf<-st_filter(access_points_sf,regional_sf)
mrt_regional_sf<-st_filter(mrt_exits_sf,regional_sf)
bikeracks_regional_sf<-st_filter(bike_racks,regional_sf)
parks_regional_sf<-st_filter(parks,regional_sf)

regional_st<-st_transform(regional_sf,crs(regional_sf))
regional_sf <- st_transform(regional_sf, crs = 6345)
regional_sp <- as_Spatial(st_zm(regional_sf))
regional_window <- as.owin(regional_sp)

```

### Regional Access Points data
```{r}
ap_regional_q <- st_transform(ap_regional_sf, crs(regional_sp))
ap_regional_ppp <- as.ppp(st_coordinates(ap_regional_q), W=regional_window)

ap_regional_Q20 <- quadratcount(ap_regional_ppp, nx=20, ny=20)
ap_regional_distribution <- intensity(ap_regional_Q20, image=TRUE)

ann_observed <- mean(nndist(ap_regional_ppp, k=1))
ann_observed
```

### Regional Amenity Density-based Monte Carlo
```{r}
amenities <- list(mrt_regional_sf, bikeracks_regional_sf, parks_regional_sf)
for (i in 1:length(amenities)) {
  amenity_regional_sf <- amenities[[i]]
  amenity_regional_q <- st_transform(amenity_regional_sf, crs(regional_sp))
  amenity_regional_ppp <- as.ppp(st_coordinates(amenity_regional_q), W=regional_window)
  
  amenity_regional_Q20 <- quadratcount(amenity_regional_ppp, nx=20, ny=20)
  amenity_regional_distribution <- intensity(amenity_regional_Q20, image=TRUE)
  
  runMC(1000, ap_regional_ppp$n, amenity_regional_distribution, ann_observed)
}
```

### Regional Amenity Distance-based Monte Carlo
```{r}
r <- raster(ncol=1000, nrow=1000, xmn= regional_sp@bbox['x','min'], xmx=regional_sp@bbox['x','max'], ymn=regional_sp@bbox['y','min'], ymx=regional_sp@bbox['y','max'])
# r <- raster(ncol=1000, nrow=1000, xmn= -700525, xmx= -694504.7, ymn= 19842439, ymx=19847018.0)

#Convert MRT exits with the same CRS as sg to an sp object
amenity_regional_q_sp <- as(amenity_regional_q, "Spatial")

#Get x and y coordinates from MRT exits
amenity_regional_distances <- distanceFromPoints(object = r, xy=coordinates(amenity_regional_q_sp)[,1:2])
plot(amenity_regional_distances)

#Mask the distances_mrt raster with the shape of Singapore
amenity_regional_distances <- mask(amenity_regional_distances, regional_sp)

#Convert distances_mrt from raster to im
amenity_regional_distances_distribution <- as.im(amenity_regional_distances)
inv_amenity_regional_distances_distribution = -amenity_regional_distances_distribution + max(amenity_regional_distances_distribution)

runMC(1000, ap_regional_ppp$n, inv_amenity_regional_distances_distribution, ann_observed)

```



# Insights

```{r}
pcn_loops <- st_read("data/park-connector-loop/park-connector-loop.kml")
unique(pcn_loops$Name)
```

```{r}
access_pts <- st_read("data/PCNs/PCN_Access_Points.geojson")
unique(access_pts$name)
```

```{r}
northern_explorer_loop <- access_pts[access_pts$name %in% c("Marsiling Park Connector", "Woodlands Park Connector", "Mandai Park Connector", "Khatib Bongsu Park Connector", "Yishun Park Connector", "Canberra-Sembawang Park Connector", "Admiralty West Park Connector", "Springleaf Park Connector", "Central Catchment Park Connector", "Canberra Park Connector", "Ulu Sembawang Park Connector", "Simpang Kiri Park Connector"),]

eastern_coastal_loop <- access_pts[access_pts$name %in% c("Bedok Park Connector", "Geylang Park Connector", "Tampines Park Connector", "Pasir Ris Park Connector", "Loyang Park Connector", "Coastal Park Connector", "Tampines-Loyang Park Connector", "Siglap Park Connector"),]

central_urban_loop <- access_pts[access_pts$name %in% c("Ang Mo Kio Avenue 5 Park Connector", "Kallang Park Connector", "Serangoon Park Connector", "Jalan Pelikat Park Connector", "Whampoa Park Connector", "Ang Mo Kio Park Connector", "Pelton Canal Park Connector", "Paya Lebar Park Connector", "Buangkok Park Connector", "Balam Park Connector"),]
  
north_eastern_riverine_loop <- access_pts[access_pts$name %in% c("Punggol Promenade Riverside Walk", "Sungei Serangoon Park Connector", "Punggol Promenade Punggol Point Walk", "Punggol Promenade Nature Walk"),]

south_ridges_loop <- access_pts[access_pts$name %in% c("Alexandra Park Connector", "Ulu Pandan Park Connector", "Pandan Gardens Park Connector", "West Coast Park Connector", "Alexandra Garden Trail", "Alexandra Canal Linear Park", "Henderson Park Connector"),]

western_adventure_loop <- access_pts[access_pts$name %in% c("Jurong Park Connector", "Jurong West Park Connector", "Choa Chu Kang North Park Connector", "Choa Chu Kang Park Connector", "Bukit Batok Park Connector", "Bukit Batok West Park Connector", "Hillview Park Connector", "Bukit Timah Park Connector", "Simei Park Connector", "Tanjong Rhu Promenade", "Bukit Panjang Park Connector", "Bukit Batok East Park Connector"),]
```

```{r}
tmap_mode("plot")

tm_shape(sg) + tm_borders() + tmap_options(check.and.fix = TRUE) + 
  tm_shape(northern_explorer_loop) + tm_symbols(size = 0.1, alpha = 0.5, col = "red") + tm_add_legend(type="symbol", labels=c("Northern Explorer Loop"), col='red') +
  tm_shape(eastern_coastal_loop) + tm_symbols(size = 0.1, alpha = 0.5, col = "darkgreen") + tm_add_legend(type="symbol", labels=c("Eastern Coastal Loop"), col='darkgreen') +
  tm_shape(central_urban_loop) + tm_symbols(size = 0.1, alpha = 0.5, col = "yellow") + tm_add_legend(type="symbol", labels=c("Central Urban Loop"), col='yellow') +
  tm_shape(north_eastern_riverine_loop) + tm_symbols(size = 0.1, alpha = 0.5, col = "purple") + tm_add_legend(type="symbol", labels=c("North Eastern Riverine Loop"), col='purple') +
  tm_shape(south_ridges_loop) + tm_symbols(size = 0.1, alpha = 0.5, col = "blue") + tm_add_legend(type="symbol", labels=c("South Ridges Loop"), col='blue') +
  tm_shape(western_adventure_loop) + tm_symbols(size = 0.1, alpha = 0.5, col = "grey") + tm_add_legend(type="symbol", labels=c("Western Adventure Loop"), col='grey') +
  tm_layout(title="Park Connector Loops", title.size = 0.65, title.position = c(0.02, "bottom"))
```

## Create buffers
```{r}
northern_explorer_loop_500buf <- st_buffer(northern_explorer_loop, dist = 500)
northern_explorer_loop_500buf <- st_union(northern_explorer_loop_500buf)

eastern_coastal_loop_500buf <- st_buffer(eastern_coastal_loop, dist = 500)
eastern_coastal_loop_500buf <- st_union(eastern_coastal_loop_500buf)

central_urban_loop_500buf <- st_buffer(central_urban_loop, dist = 500)
central_urban_loop_500buf <- st_union(central_urban_loop_500buf)

north_eastern_riverine_loop_500buf <- st_buffer(north_eastern_riverine_loop, dist = 500)
north_eastern_riverine_loop_500buf <- st_union(north_eastern_riverine_loop_500buf)

south_ridges_loop_500buf <- st_buffer(south_ridges_loop, dist = 500)
south_ridges_loop_500buf <- st_union(south_ridges_loop_500buf)

western_adventure_loop_500buf <- st_buffer(western_adventure_loop, dist = 500)
western_adventure_loop_500buf <- st_union(western_adventure_loop_500buf)
```

```{r}
loop_buffers <- list(northern_explorer_loop_500buf, eastern_coastal_loop_500buf, central_urban_loop_500buf, north_eastern_riverine_loop_500buf, south_ridges_loop_500buf, western_adventure_loop_500buf)

loop_pts <- list(northern_explorer_loop, eastern_coastal_loop, central_urban_loop, north_eastern_riverine_loop, south_ridges_loop, western_adventure_loop)
```

```{r}
dim(loop_pts[1][[1]])[1]
```

```{r}
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  print(class(loop))
}
```

## Count MRT intersection
```{r}
i = 1
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  n <- dim(loop_pts[i][[1]])[1]
  count <- poly.counts(as_Spatial(mrt_exits_sf), loop)
  print(count)
  print(count / n)
  i = i + 1
}
```

## Count bus stop intersection
```{r}
i = 1
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  n <- dim(loop_pts[i][[1]])[1]
  count <- poly.counts(spTransform(as_Spatial(busstops_sf), crs(loop)), loop)
  print(count / n)
  print(n)
  i = i + 1
}
```
## Count bike racks intersection
```{r}
i = 1
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  n <- dim(loop_pts[i][[1]])[1]
  count <- poly.counts(as_Spatial(bike_racks), loop)
  print(count / n)
  i = i + 1
}
```
## Count parks intersection
```{r}
i = 1
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  n <- dim(loop_pts[i][[1]])[1]
  count <- poly.counts(as_Spatial(parks), loop)
  print(count / n)
  i = i + 1
}
```
## Supermarkets
```{r}
i = 1
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  n <- dim(loop_pts[i][[1]])[1]
  count <- poly.counts(as_Spatial(supermarkets_sf), loop)
  print(count / n)
  i = i + 1
}
```
## Hawker centres
```{r}
i = 1
for (loop in loop_buffers) {
  loop <- as_Spatial(loop)
  n <- dim(loop_pts[i][[1]])[1]
  count <- poly.counts(as_Spatial(hawkercentres_sf), loop)
  print(count / n)
  i = i + 1
}
```



